A Control Strategy for a Tethered Follower Robot for Pulmonary Rehabilitation

### 2020-07-19 alpibot.tex
Subo una nueva version que incorpora los nuevos graficos que hizo Luciano, la explicacion de Juan sobre la equivalencia entre los mecanismos con uno y dos hilos, y varios ajustes que pidieron los revisores.

\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.  They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}. This pathology presents an musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}. In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.  During this rehabilitation, patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.   Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.  Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.  However, their own condition prevent them with the ability to precisely carry the often bulky external tank.  This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.  This scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologists and physical therapists.

@@

An alternative solution is to use an assitive ground service robot~\cite{Neto2015} to carry the oxygen tank, while following the patient in a leader-follower configuration.  There are two reasons that support the initial viability of this idea.  First,  the rehabilitation gym is a constrained environment where this problem can be tackled by an Unmanned Ground Vehicle (UGV).  On the other hand, the range of movements performed by the patient during the Rehabilitation Procedure is highly predictable by the treatment.  At the same time,  the global robotic research community looks forward for the development of robotic affordable solutions to the social and health-related worldwide problems~\cite{Khamis2019}.

@@

    % Text Node
    \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (478.5,1350) -- (535.5,1350) -- (535.5,1374) -- (478.5,1374) -- cycle  ;
    \draw (507,1362) node    {$subject$};
    % Text Node
    \draw (302,1344) node    {$left\ reel\  \ $};
    % Text Node
    \draw (299,1425) node    {$right\ reel\  \ $};
    % Text Node
    \draw (125,1414) node    {$castor\ wheel$};
    % Text Node
    \draw (422,1348) node  [rotate=-12.55]  {$left\ thread$};
    % Text Node
    \draw (423,1422) node  [rotate=-344.79]  {$right\ thread$};
    % Text Node
    \draw (348,1275) node    {$left\ wheel$};
    % Text Node
    \draw (348,1496) node    {$right\ wheel$};
    \end{tikzpicture}
    \caption{Components of the robotic vehicle and the tether mechanism}
    \label{fig:components}
\end{figure}{}

@@

% El detalle de los encoders se puede eliminar, es info común.
Encoders in each reel measure the difference in length of each thread compared to its initial position. This difference in length is the input for the control algorithm.  Using the encoder, the difference in length for each thread can be measured with Equations~\ref{eq:encoder_len}.

\begin{equation}
\label{eq:encoder_len}
\begin{array}{ll}  
l_l = pulses_{l} \; \frac{2\pi r}{ppr}  \\
\\
l_r = pulses_{r} \; \frac{2\pi r}{ppr}
\end{array}
\end{equation}{}

\noindent where pulses obtained for each reel encoder are $pulses_{l}$ and $pulses_{r}$ for the left and right encoder. These equations provide the estimated values for the left $l_l$ and right $l_r$  thread distance. The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.  The initial position of the threads is configurable.

@@

This control strategy is similar to the one presented in \cite{Ortlieb2016}. It is based on the idea that the difference in length between both threads provides a quasi-linear function of the relative angle between the subject and the vehicle orientation and, additionally, the mean of both thread length allows to approximate the relative distance between the robot and the subject.  They are described by Equations~\ref{eq:lemniscata}, \ref{eq:lemniscata2} and~\ref{eq:lemniscata3}.

\begin{equation}
v_{tar} = c_v (\frac{l_l + l_r}{2} - l_d)
\label{eq:lemniscata}
\end{equation}

\begin{equation}
\omega_{L} = v_{tar} + c_{\alpha} (l_l - l_r)
\label{eq:lemniscata2}
\end{equation}

\begin{equation}
\omega_{R} = v_{tar} - c_{\alpha} (l_l - l_r)
\label{eq:lemniscata3}
\end{equation}

\noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.  As shown in Figure~\ref{fig:follow_params},  $l_l$ and $l_r$ are thread distances obtained from encoder information from Equation~\ref{eq:encoder_len}, $v_{tar}$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the power values that are directly used to drive power to each wheel motor.

@@

\begin{equation}
\textit{if  }{\frac{l_l + l_r}{2} < l_d}\textit{ then } \omega_{L} = \omega_{R} = 0 
\end{equation}

\noindent where $l_d$ is a constant offset that is used to customize the length of the thread where the robot does not move.

@@

    % Text Node
    \draw (524,1960) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ]  {$ \begin{array}{l}
    l_{r}\\
    \end{array}$};
    % Text Node
    \draw (510,1910) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ,rotate=-358.27]  {$ \begin{array}{l}
    l_{l}\\
    \end{array}$};
    % Text Node
    \draw (496,2042) node  [color={rgb, 255:red, 128; green, 128; blue, 128 }  ,opacity=1 ]  {$subject\ start$};
    % Text Node
    \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (554.5,1904) -- (611.5,1904) -- (611.5,1928) -- (554.5,1928) -- cycle  ;
    \draw (583,1916) node    {$subject$};
    % Text Node
    \draw (390.58,1925.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$\omega_{L}$};
    % Text Node
    \draw (407.58,2100.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$\omega_{R}$};

@@

The procedure is detailed in Algorithm~\ref{alg:rotateandgoalg}.  The variable $V_r$ is the speed at which the vehicle will rotate on its axis, and $V_f$ is the speed at which the vehicle will move forward once it can move on the subject's direction.  The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_{off}$ that regulates the sensibility of the rotation movement.  The constant parameter $Dm_{off}$ is similar to $l_d$, and is used to customize the length of the thread where the robot does not move at all.

\begin{algorithm}[]
\caption{Rotate and go algorithm}
\begin{algorithmic}
\STATE $Dt \gets l_l - l_r$
\STATE $Dm \gets \frac{l_r+l_l}{2}$
\STATE $V_r \gets c_r * (abs(Dt) - Dt_{off}) + base_{vr}$ 
\STATE $V_f \gets c_v * (Dm - Dm_{off})$

\IF{$abs(Dt) > Dt_{off}$}
\IF{$l_l > l_r$}
\STATE $\omega_{L} \gets V_r$
\STATE $\omega_{R} \gets -V_r$
\ELSE
\STATE $\omega_{L} \gets -V_r$
\STATE $\omega_{R} \gets V_r$
\ENDIF

@@

\section{Equivalence between the single and double tethered system} \label{development}

@@

\draw    (351.04,132.38) .. controls (360.75,134.84) and (374.54,141.11) .. (375.89,158.8) ;
\draw [shift={(376,161.69)}, rotate = 270] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
\draw [shift={(348,131.69)}, rotate = 11.31] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;

@@

% Text Node
\draw (399,166.09) node [anchor=north west][inner sep=0.75pt]    {$x$};
% Text Node
\draw (305,12.4) node [anchor=north west][inner sep=0.75pt]    {$y$};
% Text Node
\draw (438,173.4) node [anchor=north west][inner sep=0.75pt]    {$x'$};
% Text Node
\draw (308,193.4) node [anchor=north west][inner sep=0.75pt]    {$x''$};
% Text Node
\draw (357,52.4) node [anchor=north west][inner sep=0.75pt]    {$l_{m}$};
% Text Node
\draw (439,90.4) node [anchor=north west][inner sep=0.75pt]    {$l_{R}$};
% Text Node
\draw (261,90.4) node [anchor=north west][inner sep=0.75pt]    {$l_{L}$};
% Text Node
\draw (333,114.4) node [anchor=north west][inner sep=0.75pt]    {$\theta $};
% Text Node
\draw (372,122.4) node [anchor=north west][inner sep=0.75pt]    {$\gamma $};
% Text Node
\draw (334,215.4) node [anchor=north west][inner sep=0.75pt]    {$b$};

\end{tikzpicture}
\label{fig:dibujito}
\caption{Following mechanism parameters}
\end{figure}

In~\cite{Endo2015} authors propose a similar design based on only one thread.  They propose two control strategies based on two input parameters that are obtained from a linear and circular potentiometer that determines the length of the thread $l_m$ and the orientation angle $\theta$.  We show here that they are equivalent to the approach presented based on $l_l$ and  $l_r$.

Based on a frame reference with the robot on the center of coordinates as shown in Figure~\ref{fig:dibujito}, from the input values $l_m$ and $\theta$ we can derive the position of the leader as

In this frame of reference, the lengths of the threads, $l_r$ and $l_l$ can be determined as

\begin{equation}
\begin{array}{ll}  
l_r &= \sqrt[2]{y^2 + {x'}^2} \\
l_l &= \sqrt[2]{y^2 + {x''}^2}. \\
\end{array}
\label{eq:threadsincoordinates}
\end{equation}

From Figure~\ref{fig:dibujito} we can see that ${x''} + x' =  b$. Combining this with~\ref{eq:threadsincoordinates}, we can form a system of equations to determine $x'$, $x''$ and $y$:

\begin{align*}
{x''} + x'					&=  b &&\text{(a)}  \\
{x'}^2 +  y^2                &= l_r &&\text{(b)} \\
{x''}^2 + y^2                &= l_l  &&\text{(c)}. \\
\end{align*}

From the first equation, rearranging and squaring both terms leads to

@@

\begin{align*}
{x'}^2 - {x'' }^2 = {l_r}^2 - {l_l}^2 \\
{x'}^2 - {x'' }^2 = {l_r}^2 - {l_l}^2, \\
\end{align*}

@@

\begin{align*}
b^2 - 2 b {x''}= {l_r}^2 - {l_l}^2. \\
\end{align*}

@@

\begin{align*}
x'' = \frac{({l_d}^2 - {l_i}^2 - b^2)}{-2 \; b}.    \\
\end{align*}

From Figure~\ref{fig:dibujito} it can also be seen that $\frac{b}{2}$ is equals to $x'' - x$.  Hence this can be used to finally determine $x$ and $y$ values based on the threads length $l_l$ and $l_r$ and the axle length $b$.  The $y$ value comes from Equation~\ref{eq:threadsincoordinates}.

\begin{align*}
y &= \sqrt[2]{{l_l}^2 - {x''}^2 } \\
x &= x'' - \frac{b}{2}\\
\end{align*}

@@

\begin{align*}
\sin{\gamma'} = \frac{x}{l_m}\\
\end{align*}

@@

\begin{align*}
\cos{\gamma'} = \frac{y}{l_m}\\
\end{align*}

\noindent and as $\theta = \gamma'$, we can obtain the values of $\theta$ and $l_m$:

@@

Additionally in the proposed system based on circular spring, the spring tension depends on the released thread length, hence the recovery force will vary accordingly.   The proposed scheme allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
Another issue that can be presented is when $l_m$ is equals to $l_d$.  If the patient performs a circular movement where these values are equal, the robot will not change direction at all, because $\omega_{r}$ will be also zero.   If the patient restart the movement, $\omega_{r}$ could be very low and the $K_p$ constant could not be enough to match the person movement rapidly.

Regarding the second approach, the dead-reckoning could be challenging and requires very precise measurements to reduce the error-prone characteristic of this solution~\cite{DurrantWhyte1994}.

%Otro problema que encontré es en el segundo método de control. Éste es un poco más complicado de explicar con palabras así que haré un gráfico y se los envío. Como adelanto, está en la ecuación 17 del paper que es cuando toman como medida de la diferencia de distancias (l_m - l_d) para el cálculo de la V_r.

@@

\noindent where $ \phi $ is the free parameter, and $x$ and $y$ determine the shape of the trajectory on the navigation plane.

@@

As shown in Figure~\ref{fig:capturesystem}, a tracking marker was placed on each side of the robot (on top of each thread reel). The human leader used his hand to grab the tip at which the two tethers were tied together. A third marker was placed in his hand, using a glove. The lemniscate of Gerono, used in the simulation, was marked on the floor, and the human leader tried to move his hand following this shape as close as possible, with stable speed.  The shape was marked according to the shape described in Equation~\ref{eq:lemniscata}, using $a = 2$ [m].

@@

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/capture.png}
\caption{Hardware prototype on the motion capture system and a testing subject holding the threads.  The lemniscate of gerone was marked on the floor.  The subject follows this track on the performed experiments.}
\label{fig:capturesystem}
\end{figure}

@@

\begin{center}
\begin{tabular}{ |c|c|c|c| }
\hline
$c_v$ & $c_{\alpha}$ & m.t.d. & t.t.d. \\
\hline
10  &   15  & 0.3614 & 2.0651\\
15  &   5  & 0.4325 & 2.055\\
\textbf{15}  &   \textbf{10}  & \textbf{0.2188} & \textbf{1.0902}\\
15  &   15  & 0.2891 & 1.5059\\
5  &   20  & 0.5733 & 3.7289\\
\hline
\end{tabular}
\captionof{table}{Max trajectory deviation m.t.d. [m] and Total trajectory deviation t.t.d. for different Follow the thread constants.}
\label{tab:simulationmetricsftt}
\end{center}

\begin{center}
\begin{tabular}{ |c|c|c|c|c| }
\hline
$c_v$ & $c_r$ & $Dt_{off}$ & m.t.d. & t.t.d.\\
\hline
10  &   20  &   0.1  & 0.4310 & 1.6380\\
20  &   20  &   0.05  & 0.7775 & 3.1139\\
\textbf{20}  &  \textbf{20}  & \textbf{0.1} & \textbf{0.4123} & \textbf{0.9872}\\
20  &   35  &   0.1  & 0.4143 & 1.4820\\
20  &   5  &  0.05  & 0.7815 & 3.0892\\
35  &   20  &   0.1  & 0.6337 & 1.6190\\
\hline
\end{tabular}
\captionof{table}{Max trajectory deviation m.t.d. (in [m]) and total trajectory deviation t.p.d. for different Rotate and go constants.}
\label{tab:simulationmetricsrg}
\end{center}

@@

\begin{center}
\begin{tabular}{ |c|c|c|c| }
\hline
$c_v$ & $c_{\alpha}$ & m.t.d. [m] & t.t.d. \\
\hline
\textbf{25}  &   \textbf{20}  & \textbf{0.3876} & \textbf{1.9761}\\
25  &   35  & 0.4672 & 2.3528\\
\hline
\end{tabular}
\captionof{table}{Max trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using Follow the thread.}
\label{tab:cap_ft_npd_table}
\end{center}

@@

\end{tabular}
\captionof{table}{Max trajectory deviation m.t.d. and area under normal trajectory deviation t.p.d. in motion capture experiments using Rotate and go.}
\label{tab:cap_rg_npd_table}
\end{center}

@@

>>>>

%\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.  They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}. This pathology presents a musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}. In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.  During this rehabilitation, patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.   Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activities~\cite{Celli2014}.  Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.  However, their own condition prevent them with the ability to precisely carry the often bulky external tank.  This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.  This scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologists and physical therapists.

\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.  They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}. This pathology presents a musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}. In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.  Pulmonary Rehabilitation procedures consist of controlled waking activities and physical exercises that patients perform under the supervision of a physical therapist.  However, COPD patients present a severe low oxygen saturation illness and they require oxygen supply, particularly when performing physical activities~\cite{Celli2014}.  Hence, they are required to carry an oxygen tank for the oxygenotherapy assistance, but their own condition prevent them with the ability to precisely carry the often bulky external tank.  This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.  The scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologists and physical therapists.

@@

An alternative solution is to use an assitive ground service robot~\cite{Neto2015} to carry the oxygen tank while following the patient in a leader-follower configuration.  There are two reasons that support the initial viability of this idea.  First,  the rehabilitation gym is a constrained environment where this problem can be tackled by an Unmanned Ground Vehicle (UGV).  On the other hand, the range of movements performed by the patient during the Rehabilitation Procedure is highly predictable by the treatment.  At the same time,  the global robotic research community looks forward for the development of robotic affordable solutions to social and health-related worldwide problems~\cite{Khamis2019}.

@@

    % Text Node
    \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (478.5,1350) -- (535.5,1350) -- (535.5,1374) -- (478.5,1374) -- cycle  ;
    \draw (517,1362) node    {$subject$};
    % Text Node
    \draw (302,1344) node    {$left\ reel\  \ $};
    % Text Node
    \draw (299,1425) node    {$right\ reel\  \ $};
    % Text Node
    \draw (125,1414) node    {$castor\ wheel$};
    % Text Node
    \draw (422,1348) node  [rotate=-12.55]  {$left\ thread$};
    % Text Node
    \draw (423,1422) node  [rotate=-344.79]  {$right\ thread$};
    % Text Node
    \draw (348,1275) node    {$left\ wheel$};
    % Text Node
    \draw (348,1496) node    {$right\ wheel$};
    \end{tikzpicture}
    \caption{Components of the robotic vehicle and the tether mechanism}
    \label{fig:components}
\end{figure}{}

@@

% El detalle de los encoders se puede eliminar, es info común.
Encoders in each reel measure the difference in length of each thread compared to its initial position. These differences in length are the input for the control algorithm.  Using the encoder, the difference in length for each thread can be measured with Equations~\ref{eq:encoder_len}.

\begin{equation}
\label{eq:encoder_len}
\begin{array}{ll}  
l_L = pulses_{l} \; \frac{2\pi r}{ppr}  \\
\\
l_R = pulses_{r} \; \frac{2\pi r}{ppr}
\end{array}
\end{equation}{}

\noindent where pulses obtained for each reel encoder are $pulses_{l}$ and $pulses_{r}$ for the left and right encoder. These equations provide the estimated values for the left $l_L$ and right $l_R$  thread distance. The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.  The initial position of the threads is configurable.

@@

This control strategy is similar to the one presented in \cite{Ortlieb2016}.  It is based on the idea that both the relative angle between the subject and the vehicle orientation and, the relative distance between the robot and the subject, can be computed from the length of the left and right threads.  They are described by Equations~\ref{eq:lemniscata}, \ref{eq:lemniscata2} and~\ref{eq:lemniscata3}.

\begin{equation}
V_t = c_v (\frac{l_L + l_R}{2} - l_D)
\label{eq:lemniscata}
\end{equation}

\begin{equation}
\omega_{L} = V_t + c_{\alpha} (l_L - l_R)
\label{eq:lemniscata2}
\end{equation}

\begin{equation}
\omega_{R} = V_t - c_{\alpha} (l_L - l_R)
\label{eq:lemniscata3}
\end{equation}

\noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration whose units are expressed in $[\frac{1}{s}]$, and $l_D$ is a constant offset in $[m]$ that is used to customize the desired length of the thread where the robot does not move.   As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are thread distances in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.

@@

\begin{equation}
\textit{if  }{\frac{l_L + l_R}{2} < l_D}\textit{ then } \omega_{L} = \omega_{R} = 0.
\end{equation}

@@

    % Text Node
    \draw (524,1960) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ]  {$ \begin{array}{l}
    l_{R}\\
    \end{array}$};
    % Text Node
    \draw (510,1910) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ,rotate=-358.27]  {$ \begin{array}{l}
    l_{L}\\
    \end{array}$};
    % Text Node
    \draw (496,2042) node  [color={rgb, 255:red, 128; green, 128; blue, 128 }  ,opacity=1 ]  {$subject\ start$};
    % Text Node
    \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (554.5,1904) -- (611.5,1904) -- (611.5,1928) -- (554.5,1928) -- cycle  ;
    \draw (583,1916) node    {$subject$};
    % Text Node
    \draw (390.58,1925.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$\omega_{L}$};
    % Text Node
    \draw (407.58,2100.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$\omega_{R}$};

@@

The procedure is detailed in Algorithm~\ref{alg:rotateandgoalg}.  The variable $V_r$ is the speed at which the vehicle will rotate on its axis, and $V_f$ is the speed at which the vehicle will move forward once it can move on the subject's direction.  The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_{off}$ that regulates the sensibility of the rotation movement.  The constant parameter $Dm_{off}$ is similar to $l_D$, and is used to customize the length of the thread where the robot does not move at all.

\begin{algorithm}[]
\caption{Rotate and go algorithm}
\begin{algorithmic}
\STATE $Dt \gets l_L - l_R$
\STATE $Dm \gets \frac{l_R+l_L}{2}$
\STATE $V_r \gets c_r * (abs(Dt) - Dt_{off}) + base_{vr}$ 
\STATE $V_f \gets c_v * (Dm - Dm_{off})$

\IF{$abs(Dt) > Dt_{off}$}
\IF{$l_L > l_R$}
\STATE $\omega_{L} \gets V_r$
\STATE $\omega_{R} \gets -V_r$
\ELSE
\STATE $\omega_{L} \gets -V_r$
\STATE $\omega_{R} \gets V_r$
\ENDIF

@@

\section{Comparision with alternative methods} \label{development}

\subsection{Equivalence between the single and double tethered system}

In~\cite{Endo2015} authors propose a similar design based on only one thread.  They propose two control strategies based on two input parameters that are obtained from a linear and circular potentiometer that determines the length of the thread $l_m$ and the orientation angle $\theta$.  We show here that they are equivalent to the approach presented based on $l_L$ and  $l_R$.

Based on a frame reference with the robot on the center of coordinates as shown on Figure~\ref{fig:dibujitos}, from the input values $l_m$ and $\theta$ we can derive the position of the leader as

In this frame of reference, the lengths of the threads, $l_R$ and $l_L$ can be determined as

@@

\draw    (384.56,66.54) .. controls (391.15,78.31) and (389.42,81.22) .. (403.17,79.98) ;
\draw [shift={(406,79.69)}, rotate = 533.6600000000001] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
\draw [shift={(382.95,63.79)}, rotate = 58.73] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;

@@

%Shape: Brace [id:dp4829292799998628] 
\draw   (409,183.69) .. controls (409,188.36) and (411.33,190.69) .. (416,190.69) -- (418.5,190.69) .. controls (425.17,190.69) and (428.5,193.02) .. (428.5,197.69) .. controls (428.5,193.02) and (431.83,190.69) .. (438.5,190.69)(435.5,190.69) -- (441,190.69) .. controls (445.67,190.69) and (448,188.36) .. (448,183.69) ;
%Shape: Circle [id:dp5543776560649496] 
\draw   (330.5,13.69) .. controls (330.5,12.31) and (331.62,11.19) .. (333,11.19) .. controls (334.38,11.19) and (335.5,12.31) .. (335.5,13.69) .. controls (335.5,15.07) and (334.38,16.19) .. (333,16.19) .. controls (331.62,16.19) and (330.5,15.07) .. (330.5,13.69) -- cycle ;

% Text Node
\draw (399,166.09) node [anchor=north west][inner sep=0.75pt]    {$x$};
% Text Node
\draw (312,2.4) node [anchor=north west][inner sep=0.75pt]    {$y$};
% Text Node
\draw (422,194.4) node [anchor=north west][inner sep=0.75pt]    {$x'$};
% Text Node
\draw (308,193.4) node [anchor=north west][inner sep=0.75pt]    {$x''$};
% Text Node
\draw (357,52.4) node [anchor=north west][inner sep=0.75pt]    {$l_{m}$};
% Text Node
\draw (439,90.4) node [anchor=north west][inner sep=0.75pt]    {$l_{R}$};
% Text Node
\draw (261,90.4) node [anchor=north west][inner sep=0.75pt]    {$l_{L}$};
% Text Node
\draw (342,86.4) node [anchor=north west][inner sep=0.75pt]    {$\theta $};
% Text Node
\draw (379,76.4) node [anchor=north west][inner sep=0.75pt]    {$\gamma $};
% Text Node
\draw (334,215.4) node [anchor=north west][inner sep=0.75pt]    {$b$};

\end{tikzpicture}

\caption{The center of coordinates is on the robot center of mass, where the y positive axis points in the same direction as the robot direction.}
\label{fig:dibujitos}
\end{figure}{}

\begin{equation}
\begin{array}{ll}  
l_R &= \sqrt[2]{y^2 + {x'}^2} \\
l_L &= \sqrt[2]{y^2 + {x''}^2}. \\
\end{array}
\label{eq:threadsincoordinates}
\end{equation}

From Figure~\ref{fig:dibujitos} we can see that ${x''} + x' =  b$, with $b$ the axle length. Combining this with~\ref{eq:threadsincoordinates}, we can form a system of equations to determine $x'$, $x''$ and $y$:

\begin{align*}
{x''} + x'					&=  b &&\text{(a)}  \\
{x'}^2 +  y^2                &= {l_R}^2 &&\text{(b)} \\
{x''}^2 + y^2                &= {l_L}^2  &&\text{(c)}. \\
\end{align*}

From the first equation (a), rearranging and squaring both terms leads to

@@

\begin{align*}
{x'}^2 - {x'' }^2 = {l_R}^2 - {l_L}^2 \\
\end{align*}

@@

\begin{align*}
b^2 - 2 b {x''}= {l_R}^2 - {l_L}^2. \\
\end{align*}

@@

\begin{align*}
x'' = \frac{({l_R}^2 - {l_L}^2 - b^2)}{-2 \; b}.    \\
\end{align*}

From Figure~\ref{fig:dibujitos} it can also be seen that $\frac{b}{2}$ is equals to $x'' - x$.  Hence this can be used to finally determine $x$ and $y$ values based on the threads length $l_L$ and $l_R$ and the axle length $b$.  The $y$ value comes from Equation~\ref{eq:threadsincoordinates}.

\begin{align*}
y &= \sqrt[2]{{l_L}^2 - {x''}^2 } \\
x &= x'' - \frac{b}{2}\\
\end{align*}

@@

\begin{align*}
\sin{\gamma} = \frac{x}{l_m}\\
\end{align*}

@@

\begin{align*}
\cos{\gamma} = \frac{y}{l_m}\\
\end{align*}

\noindent and as $\theta = \gamma$, we can obtain the values of $\theta$ and $l_m$:

@@

\subsection{Design comparison}

In the scheme proposed by~\cite{Endo2015} a single thread is recovered mechanically by means of a circular flat spring. This device works intensively when the robot is following the patient and, therefore, the spring is exposed to wear. Instead, the double thread recovery implemented by an active reel spring allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.

Additionally with a circular flat spring, the spring tension depends on the released thread length, hence the recovery force will vary accordingly. The scheme based on an actuated reel allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.

Regarding the control algorithm, authors in \cite{Endo2015} introduced two methods. The first of them computes the angular velocity as a  function of the difference between the measured thread length $l_m$ and the desired distance to patient  $l_D$. In this way, if the patient moves around the robot with a $l_m$ equals to $l_d$, the robot will not adjust its direction until the patient stops turning and continues moving forward. Hence, the robot must correct its direction but the correction angle could be large, which leads the robot to deviate off the desired trajectory. Additionally, the second method proposed is based on dead-reckoning to estimate the patient position. This strategy requires very precise measurements to reduce the cumulative error, which is characteristic of this estimation method~\cite{DurrantWhyte1994}.

%Another issue that can be presented is when $l_m$ is equals to $l_D$.  If the patient performs a circular movement where these values are equal, the robot will not change direction at all, because $\omega_{r}$ will be also zero.   If the patient restart the movement, $\omega_{r}$ could be very low and the proportional gain $K_p$ could not be enough to match the person movement rapidly.

%Additionally, authors in~\cite{Endo2015} propose an alternative control strategy based on dead-reckoning requires very precise measurements to reduce the error-prone characteristic of this solution~\cite{DurrantWhyte1994}.

%Otro problema que encontré es en el segundo método de control. Éste es un poco más complicado de explicar con palabras así que haré un gráfico y se los envío. Como adelanto, está en la ecuación 17 del paper que es cuando toman como medida de la diferencia de distancias (l_m - l_D) para el cálculo de la V_r.

@@

\noindent where $ \phi $ is the free parameter, $a$ are the limit length of the arc of the curve, and $x$ and $y$ are the parametric functions that determine the shape of the trajectory on the navigation plane.

@@

As shown on Figure~\ref{fig:capturesystem}, a tracking marker was placed on each side of the robot (on top of each thread reel). The human leader used his hand to grab the tip at which the two tethers were tied together. A third marker was placed in his hand, using a glove. The lemniscate of Gerono, used in the simulation, was marked on the floor, and the human leader tried to move his hand following this shape as close as possible, with stable speed.  The shape was marked according to the shape described in Equation~\ref{eq:lemniscata}, using $a = 2$ [m].

@@

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/capture.png}
\caption{Hardware prototype on the motion capture system and a testing subject holding the threads.  The lemniscate of Gerono was marked on the floor.  The subject follows this track on the performed experiments.}
\label{fig:capturesystem}
\end{figure}

@@

\begin{center}
\begin{tabular}{ |c|c|c|c| }
\hline
$c_v$ & $c_{\alpha}$ & m.t.d. & t.t.d. \\
\hline
10  &   15  & 0.3614 & 2.0651\\
15  &   5  & 0.4325 & 2.055\\
\textbf{15}  &   \textbf{10}  & \textbf{0.2188} & \textbf{1.0902}\\
15  &   15  & 0.2891 & 1.5059\\
5  &   20  & 0.5733 & 3.7289\\
\hline
\end{tabular}
\captionof{table}{Maximum trajectory deviation m.t.d. [m] and Total trajectory deviation t.t.d. for different Follow the thread constants.}
\label{tab:simulationmetricsftt}
\end{center}

\begin{center}
\begin{tabular}{ |c|c|c|c|c| }
\hline
$c_v$ & $c_r$ & $Dt_{off}$ & m.t.d. & t.t.d.\\
\hline
10  &   20  &   0.1  & 0.4310 & 1.6380\\
20  &   20  &   0.05  & 0.7775 & 3.1139\\
\textbf{20}  &  \textbf{20}  & \textbf{0.1} & \textbf{0.4123} & \textbf{0.9872}\\
20  &   35  &   0.1  & 0.4143 & 1.4820\\
20  &   5  &  0.05  & 0.7815 & 3.0892\\
35  &   20  &   0.1  & 0.6337 & 1.6190\\
\hline
\end{tabular}
\captionof{table}{Maximum trajectory deviation m.t.d. (in [m]) and total trajectory deviation t.p.d. for different Rotate and go constants.}
\label{tab:simulationmetricsrg}
\end{center}

@@

\begin{center}
\begin{tabular}{ |c|c|c|c| }
\hline
$c_v$ & $c_{\alpha}$ & m.t.d. [m] & t.t.d. \\
\hline
\textbf{25}  &   \textbf{20}  & \textbf{0.3876} & \textbf{1.9761}\\
25  &   35  & 0.4672 & 2.3528\\
\hline
\end{tabular}
\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using Follow the thread.}
\label{tab:cap_ft_npd_table}
\end{center}

@@

\end{tabular}
\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.p.d. in motion capture experiments using Rotate and go.}
\label{tab:cap_rg_npd_table}
\end{center}

@@

\subsection{Validation carrying the oxygen tank}
